Java Sorting Algorithms Comparison

### 2015-02-23 Programming Project B/paper.tex
'finished ' paper

%% BEGIN ABSTRACT
\begin{abstract}

%% END ABSTRACT
\end{abstract}

@@

\textbf{Merge Sort Conclusion}: %%TODO: Write this

@@

\textbf{Merge Together Conclusion}: %%TODO: Write this.

A heap sort can be implemented in a variety of languages using the pseudocode below in Algorithm 3.
\newline
\textbf{Heap Sort Pre-Condition}: A is a non-empty array with a natural order.
\newline
\textbf{Heap Sort Post-Condition}: A' is a permutation of A in strictly non-decreasing order.
\newline
\textbf{Heapify Pre-Condition}: A is a non-empty array with a comparable data type in a natural order.
\newline
\textbf{Heapify Post-Condition}: %%TODO: Write this.
\newline
\begin{algorithm}
\caption {\textsc{Heap-Sort}(A)}
\label{algo:heapsort}
\begin{algorithmic}[1]
\Procedure{Heapify}{A, i, total}
\State{$var$ $left = i * 2$}
\State{$var$ $right = left + 1$}
\State{$var$ $iPrime = i$}
\If{$left <= total$ and $A[left] > A[i]$}
	\State{$i = left$}
\EndIf
\If{$right <= total$ and $A[right] > A[i]$}
	\State{$i = right$}
\EndIf
\If{$i != iPrime$}
	\State{$var$ $temp = A[iPrime]$}
	\State{$A[iPrime] = A[i]$}
	\State{$A[i] = tmp$}
	\State{$A = heapify(A, i, total)$}
\EndIf
\State{\Return{$A$}}
\EndProcedure
\Procedure{Heap-Sort}{A}
\State{$var$ $size = A.length$}
\For{$var$ $i = size / 2$ downto $1$}
	\State{$A = heapify(A, i, size$}
\EndFor
\For{$var$ $i = size$ downto $1$}
	\State{$var$ $tmp = A[1]$}
	\State{$A[0] = A[i]$}
	\State{$A[i] = tmp$}
	\State{$size = size - 1$}
	\State{$A = heapify(A, 1, size)$}
\EndFor
\State{\Return{$A$}}
\EndProcedure
\end{algorithmic}
\end{algorithm}
\newline
\textbf{Heap Sort First Loop Invariant}: A[i] is the parent element in a heap.
\newline
\textbf{Heap Sort First For Loop Invariant Initialization}: The invariant holds true before the first iteration of the loop because i is less than the size of the array, and therefore has a child node.
\newline
\textbf{Heap Sort First For Loop Invariant Maintenance}: The invariant holds true at the beginning of each iteration of the loop because A[i] will always be smaller than the size of the array, and therefore have children nodes.
\newline
\textbf{Heap Sort First For Loop Invariant Termination}: The invariant holds true at the termination of the loop because i will be the smallest index value of A and therefore will have children nodes.
\newline
\newline
%% TODO: something about a conclusion
\textbf{Heap Sort Second For Loop Invariant}: All elements in A at the index value greater than i are in stricly non-decreasing order.
\newline
\textbf{Heap Sort Second For Loop Invariant Initialization}: The invariant holds vacuously true before the first iteration of the loop because there are no elements in A that are at an index value greater than i.
\newline
\textbf{Heap Sort Second For Loop Invariant Maintenance}: The invariant holds true at the beginning of each iteration of the loop because each most extreme element is moved to the very end of the array at an index value greater than i.
\newline
\textbf{Heap Sort Second For Loop Invariant Termination}: The invariant holds true at the termination of the loop because i decreases as ecah largest element is moved to the end of A until the entire array has been traversed, so that all elements greater than i are in stirctly non-decreasing order.
\newline
\newline
%% TODO: Something about a conclusion here or some shit like that
%% END PROCEDURE

@@

All arrays used in testing are Java ArrayList<Integer> unless otherwise spcified. All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgewock and Kevin Wayne %%TODO: Cite this shit.

>>>>

%% BEGIN ABSTRACT
\begin{abstract}
The present study investiages the insertion sort, merge sort, and heap sort sorting algorithms and how they perform on data sets of different sizes and types whe implemented in Java. Insertion sort is a sorting algorithm that runs in $ \theta(n^2) $. For sorting large sets of data, insertion sort does not perform exemplary, however, for simple sorting of data sets under 10,000, insertion sort performs resonably quickly on modern day computers. Merge sort and heap sort both perform at $ \theta(n*\lg(n)) $. Both merge and heap sort perform exceptionally well when sorting both small and large sets of data, even upto 10,000,000 elements.
%% END ABSTRACT
\end{abstract}

@@

@@

A heap sort can be implemented in a variety of languages using the pseudocode below in Algorithm 3.
\newline
\textbf{Heap Sort Pre-Condition}: A is a non-empty array with a natural order.
\newline
\textbf{Heap Sort Post-Condition}: A' is a permutation of A in strictly non-decreasing order.
\newline
\textbf{Heapify Pre-Condition}: A is a non-empty array with a comparable data type in a natural order.
\newline
\textbf{Heapify Post-Condition}: Each parent node is more extreme than its child  node.
\newline
\begin{algorithm}
\caption {\textsc{Heap-Sort}(A)}
\label{algo:heapsort}
\begin{algorithmic}[1]
\Procedure{Heapify}{A, i, total}
\State{$var$ $left = i * 2$}
\State{$var$ $right = left + 1$}
\State{$var$ $iPrime = i$}
\If{$left <= total$ and $A[left] > A[i]$}
	\State{$i = left$}
\EndIf
\If{$right <= total$ and $A[right] > A[i]$}
	\State{$i = right$}
\EndIf
\If{$i != iPrime$}
	\State{$var$ $temp = A[iPrime]$}
	\State{$A[iPrime] = A[i]$}
	\State{$A[i] = tmp$}
	\State{$A = heapify(A, i, total)$}
\EndIf
\State{\Return{$A$}}
\EndProcedure
\Procedure{Heap-Sort}{A}
\State{$var$ $size = A.length$}
\For{$var$ $i = size / 2$ downto $1$}
	\State{$A = heapify(A, i, size$}
\EndFor
\For{$var$ $i = size$ downto $1$}
	\State{$var$ $tmp = A[1]$}
	\State{$A[0] = A[i]$}
	\State{$A[i] = tmp$}
	\State{$size = size - 1$}
	\State{$A = heapify(A, 1, size)$}
\EndFor
\State{\Return{$A$}}
\EndProcedure
\end{algorithmic}
\end{algorithm}
\newline
\textbf{Heap Sort First Loop Invariant}: A[i] is the parent element in a heap.
\newline
\textbf{Heap Sort First For Loop Invariant Initialization}: The invariant holds true before the first iteration of the loop because i is less than the size of the array, and therefore has a child node.
\newline
\textbf{Heap Sort First For Loop Invariant Maintenance}: The invariant holds true at the beginning of each iteration of the loop because A[i] will always be smaller than the size of the array, and therefore have children nodes.
\newline
\textbf{Heap Sort First For Loop Invariant Termination}: The invariant holds true at the termination of the loop because i will be the smallest index value of A and therefore will have children nodes.
\newline
\newline
\textbf{Heap Sort Second For Loop Invariant}: All elements in A at the index value greater than i are in stricly non-decreasing order.
\newline
\textbf{Heap Sort Second For Loop Invariant Initialization}: The invariant holds vacuously true before the first iteration of the loop because there are no elements in A that are at an index value greater than i.
\newline
\textbf{Heap Sort Second For Loop Invariant Maintenance}: The invariant holds true at the beginning of each iteration of the loop because each most extreme element is moved to the very end of the array at an index value greater than i.
\newline
\textbf{Heap Sort Second For Loop Invariant Termination}: The invariant holds true at the termination of the loop because i decreases as ecah largest element is moved to the end of A until the entire array has been traversed, so that all elements greater than i are in stirctly non-decreasing order.
\newline
\newline
%% END PROCEDURE

@@

All arrays used in testing are Java ArrayList<Integer> unless otherwise spcified. All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgewock and Kevin Wayne \cite{stopwatch}
